Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-03-31 Grassmannian_cells.tex
First pass on representation theory of (n)$: truncated projectives

@@

\renewcommand{\AA}{\mathbb{A}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}

@@

@@

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.  For $m\ge1$, let $P_m$ be the preprojective representation of $K(n)$ with dimension vector $(u_m,u_{m-1})$.

>>>>

\input xy
\xyoption{all}

@@

\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Representation Theory of $K(n)$}

Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K(n)_0=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$.

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
\begin{theorem}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively.
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}P_{m+1}^{a_2}$ or $I_m^{a_1}I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.

\begin{lemma}
  For any $m\ge1$, we have $\dim\Hom(P_m,P_m)=1$, $\dim\Hom(P_m,P_{m+1})=n$, and $\dim\Hom(P_{m+1},P_m)=0$.  
  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
\end{lemma}
Write $P_{m+1}^{\{f_1,\ldots,f_k\}}$ for the cokernel of the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$, i.e. we have a short exact sequence
\begin{equation}
\label{eq:truncated preprojectives}
  0\longrightarrow P_m^k\stackrel{[f_1\ \cdots\ f_k]}{\longrightarrow} P_{m+1}\longrightarrow P_{m+1}^{\{f_1,\ldots,f_k\}}\longrightarrow 0.
\end{equation}
\begin{lemma}
  For $m\ge1$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the following hold:
  \begin{enumerate}
    \item $\Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})$ is naturally isomorphic to $\Hom(P_m,P_{m+1})/\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ is naturally isomorphic to $\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Hom(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)=0$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Applying the functor $\Hom(P_m,-)$ to the sequence \eqref{eq:truncated preprojectives} gives the exact sequence
  \[0\longrightarrow \Hom(P_m,P_m^k)\stackrel{[f_1\ \cdots\ f_k]\circ}{\longrightarrow} \Hom(P_m,P_{m+1})\longrightarrow \Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})\longrightarrow 0,\]
  establishing (1).

  To see (2), we apply the functor $\Hom(-,P_m)$ to the sequence \eqref{eq:truncated preprojectives} to get an isomorphism $\Hom(P_m^k,P_m)\cong\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$.
  To get the isomorphism with $\kk f_1\oplus\cdots\oplus\kk f_k$, we work by induction on $k$.  
  Indeed, $\Hom(P_m^k,P_m)$ is spanned by the coordinate projections $\pi_i:P_m^k\to P_m$ for $1\le i\le k$ and each of these gives rise to a commutative diagram
  \[\xymatrix{ & 0\ar[d] & 0\ar[d] & & \\ & P_m^{k-1}\ar[d]\ar@{=}[r]& P_m^{k-1}\ar[d]^{[f_1\ \cdots\ \widehat{f_i}\ \cdots\ f_k]}& & \\
    0\ar[r]& P_m^k\ar[r]^{[f_1\ \cdots\ f_k]}\ar[d]^{\pi_i}\ar@{}[dr]|{\ulcorner}& P_{m+1}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]\ar@{=}[d]& 0 \\
    0\ar[r]& P_m\ar[r]^(.35){f_i}\ar[d]& P_{m+1}^{\{f_1,\ldots,\widehat{f_i},\ldots,f_k\}}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]& 0\\
     & 0 & 0 & & }\]
  The bottom row of this diagram is the extension in $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ naturally identified with the generator $f_i$ of $\kk f_1\oplus\cdots\oplus\kk f_k$.

  Part (3) follows immediately from the observation that $\langle P_{m+1}^{\{f_1,\ldots,f_k\}},P_m\rangle=-k$.
\end{proof}

@@
